Predicting trends in the quality of state-of-the-art neural networks without access to training or testing data

### 2020-12-31 submission/NatCom_all_models.tex
getting appendix in place
pass on everything except nlp subsection

@@

\begin{table}[t]
\small
\begin{center}
\begin{tabular}{|p{1in}|c|c|c|c|}
\hline
%    & Frobenius Norm & Spectral Norm & Weighted Alpha & Alpha-Norm \\
%    & $\langle\log\Vert\mathbf{W}\Vert_{F}\rangle$ & $\langle\log\Vert\mathbf{W}\Vert_{\infty}\rangle$ & $\langle\hat{\alpha}=\alpha\log\lambda_{max}\rangle$ & $\langle\log\Vert\mathbf{X}\Vert^{\alpha}_{\alpha}\rangle$ \\
Series        & $\log\Vert\cdot\Vert_{F}$ & $\log\Vert\cdot\Vert_{\infty}$ & $\hat{\alpha}$ & $\log\Vert\cdot\Vert^{\alpha}_{\alpha}$ \\
\hline
$R^{2}$ (mean) & 0.63 &  0.55 & \textbf{0.64} & \textbf{0.64} \\
$R^{2}$ (std)  & 0.34 &  0.36 & \textbf{0.29} &          0.30 \\
\hline
$MSE$ (mean)   & 4.54 &  9.62 &          3.14 & \textbf{2.92} \\
$MSE$ (std)    & 8.69 & 23.06 &          5.14 & \textbf{5.00} \\
\hline
\end{tabular}
\end{center}
\caption{Comparison of linear regression fits for different average Log Norm and Weighted Alpha metrics across 5 CV datasets, 17 architectures, covering 108 (out of over 400) different pretrained DNNs.  
         We include regressions only for architectures with five or more data points, and which are positively correlated with test error.
         These results can be readily reproduced using the Google Colab notebooks. 
         (See the Supplementary Information
         %~\ref{sxn:appendix} 
         for details.) 
        }
\label{table:results}
\end{table}

>>>>

\michael{Refine text in this subsection after we get the final version of all the tables and figures in the appendix.  Things to emphasize: we have a huge amount of data that we are summarizing very succicently; the table here is the tip of the iceberg; the results show our metrics are good, as a sanity check; they show that you can get some results in general, but looking at the details in the appendix show that results are more reliable within a given architecture series than across different architectures; point to out publicly-availab e repo; mention that we do Kendall-$\tau$ rank correlation also.}

@@

\begin{table}[t]
\small
\begin{center}
\begin{tabular}{|p{1.25in}|c|c|c|c|}
\hline
%    & Frobenius Norm & Spectral Norm & Weighted Alpha & Alpha-Norm \\
%    & $\langle\log\Vert\mathbf{W}\Vert_{F}\rangle$ & $\langle\log\Vert\mathbf{W}\Vert_{\infty}\rangle$ & $\langle\hat{\alpha}=\alpha\log\lambda_{max}\rangle$ & $\langle\log\Vert\mathbf{X}\Vert^{\alpha}_{\alpha}\rangle$ \\
                        & $\log\Vert\cdot\Vert_{F}$ & $\log\Vert\cdot\Vert_{\infty}$ & $\hat{\alpha}$ & $\log\Vert\cdot\Vert^{\alpha}_{\alpha}$ \\
\hline
$R^{2}$ (mean)          & 0.63 &  0.55 & \textbf{0.64} & \textbf{0.64} \\
$R^{2}$ (std)           & 0.34 &  0.36 & \textbf{0.29} &          0.30 \\
\hline
$MSE$ (mean)            & 4.54 &  9.62 &          3.14 & \textbf{2.92} \\
$MSE$ (std)             & 8.69 & 23.06 &          5.14 & \textbf{5.00} \\
\hline
Kendall-$\tau$ (mean)   &  XXX &   XXX &           XXX &           XXX \\
Kendall-$\tau$ (std)    &  XXX &   XXX &           XXX &           XXX \\
\hline
\end{tabular}
\end{center}
\caption{Comparison of linear regression fits for different average Log Norm and Weighted Alpha metrics across 5 CV datasets, 17 architectures, covering 108 (out of over 400) different pretrained DNNs.  
         We include regressions only for architectures with five or more data points, and which are positively correlated with test error.
         These results can be readily reproduced using the Google Colab notebooks. 
         (See the Supplementary Information
         %~\ref{sxn:appendix} 
         for details.) 
        }
\label{table:results}
\end{table}

### 2020-12-31 submission/NatCom_discussion.tex
getting appendix in place
pass on everything except nlp subsection

Our work, starting in 2018 with the \texttt{WeightWatcher} tool~\cite{weightwatcher_package}, is the first to perform a detailed analysis of the weight matrices of DNNs~\cite{MM18_TR, MM19_HTSR_ICML, MM20_SDM}.
Subsequent to the initial version of this paper, we became aware of two other works that were posted to the in 2020 within weeks of the initial version of this paper~\cite{EJRUY20_TR,UKGBT20_TR}.
Both of these papers validate our basic result that one can gain substantial insight into model quality by examining weight matrices without access to any training or testing data.
However, both consider smaller models drawn from a much narrower range of applications than we consider.
Previous results in HT-SR Theory suggest that insights from these smaller models may not extend to the state-of-the-art CV and NLP models we consider.

>>>>

Our work also has intriguing similarities and differences with work on understanding DNNs with the information bottleneck principle~\cite{TZ15,ST17_TR}, which posits that DNNs can be quantified by the mutual information between their layers and the input and output variables.
Most importantly, our approach does not require access to any data, while information measures used in the information bottleneck approach do require this.
Nevertheless, several results from HT-SR Theory, on which our metrics are based, have parallels in the information bottleneck approach.
Perhaps most notably, the quick transition from a \textsc{Random-like} phase to \textsc{Bulk+Spikes} phase, followed by slow transition to a \textsc{Heavy-Tailed} phase, as noted previously~\cite{MM18_TR}, is reminiscent of the dynamics on the Information Plane~\cite{ST17_TR}.

Finally, 
our work, starting in 2018 with the \texttt{WeightWatcher} tool~\cite{weightwatcher_package}, is the first to perform a detailed analysis of the weight matrices of DNNs~\cite{MM18_TR, MM19_HTSR_ICML, MM20_SDM}.
Subsequent to the initial version of this paper, we became aware of two other works that were posted to the in 2020 within weeks of the initial version of this paper~\cite{EJRUY20_TR,UKGBT20_TR}.
Both of these papers validate our basic result that one can gain substantial insight into model quality by examining weight matrices without access to any training or testing data.
However, both consider smaller models drawn from a much narrower range of applications than we consider.
Previous results in HT-SR Theory suggest that insights from these smaller models may not extend to the state-of-the-art CV and NLP models we consider.
